---
title: "EstimateFossilLMA"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data Prep
Loading require packages
```{r}
require(BIEN)
require(plyr)
require(tidyr)
require(dplyr)
require(mosaic)
require(stringr)
require(lme4)
require(magrittr)
```

## Mixed-Effects Modeling

First, we need to create the fossil data for prediction.
```{r}
florissant_fossil_int <- readRDS("~/Documents/BEIN data R/data/processed/04_florissant_fossil_clean.rds")
renova_fossil_int <- readRDS("~/Documents/BEIN data R/data/processed/04_renova_fossil_clean.rds")
bridgecreek_fossil_int <- readRDS("~/Documents/BEIN data R/data/processed/04_bridgecreek_fossil_clean.rds")

fossil_comb <- rbind(florissant_fossil_int,renova_fossil_int,bridgecreek_fossil_int)
fossil_gen <- fossil_comb$Genus
fossil_gen <- unique(fossil_gen)
fossil_gen_df <- as.data.frame(fossil_gen)
fossil_tax <- BIEN_taxonomy_genus(fossil_gen)
fossil_tax <- fossil_tax[-c(1,7:9)]
fossil_tax <- unique(fossil_tax)

all_fossil <- rbind(florissant_fossil_int, renova_fossil_int, bridgecreek_fossil_int)
all_fossil$binomial <- paste(all_fossil$Genus, all_fossil$species)
all_fossil$binomial <- str_replace_all(all_fossil$binomial,"\\s+","_")
all_fossil <- all_fossil[-c(1,2)]

 all_fossil_LMEpred <- all_fossil %>%
  separate(binomial, 
    c("scrubbed_genus", "species"))
all_fossil_royer_pred <- left_join(all_fossil_LMEpred, fossil_tax, by = "scrubbed_genus")

all_fossil_royer_pred <- na.omit(all_fossil_royer_pred)
all_fossil_royer_pred <- unique(all_fossil_royer_pred)

colnames(all_fossil)[colnames(all_fossil)=="Petiole Width (cm)"] <- "avg_petiole_width"
colnames(all_fossil)[colnames(all_fossil)=="Leaf Area (cm^2)"] <- "avg_LA"
colnames(all_fossil)[colnames(all_fossil)=="PW^2/A"] <- "log_pet_leafarea"
colnames(all_fossil)[colnames(all_fossil)=="LMA (g/m^2)"] <- "log_LMA"
```

Next, we have to apply a fit linear mixed-effects model. By using the lmer function in the lme4 package, we are able to fit the model to the selected data. 
```{r mixedeffects}
royer_tax_full <- readRDS("~/Documents/BEIN data R/data/processed/07_lm4_royer")
royer_tax_na_omit<-na.omit(royer_tax_full)

royer_lme <- lmer(log_lma~royer_tax_full$log_pet_leafarea + (1|order/scrubbed_family/scrubbed_genus), data=royer_tax_full)
royer_lme_sum<- summary(royer_lme)


royer_pred <- as.data.frame(predict(royer_lme, newdata = all_fossil_royer_pred, allow.new.levels= TRUE))

#lm predictions and models using family as factor
royer_fossil_dropped <-   all_fossil_royer_pred %>% 
  filter(!grepl('Cercidiphyllaceae', scrubbed_family)) %>%
  filter(!grepl('Smilacaceae', scrubbed_family)) %>%
  filter(!grepl('Staphyleaceae', scrubbed_family))

royer_fossil_dropped <-   all_fossil_royer_pred %>% 
  filter(!grepl('Cercidiphyllaceae', scrubbed_family)) %>%
  filter(!grepl('Smilacaceae', scrubbed_family)) %>%
  filter(!grepl('Staphyleaceae', scrubbed_family))

royer_tax_fam_count <- royer_tax_full%>% 
  group_by(scrubbed_family) %>% 
  tally() %>%
  filter(!grepl('Unknown', scrubbed_family))

royer_tax_count <- left_join(royer_tax_full, royer_tax_fam_count, by="scrubbed_family")

```

###Preparing Logistic Regression
Determining the top 5 extant families with the most samples, and the top 10 families respectively. 
```{r}
royer_tax_top5 <- subset(royer_tax_count, as.numeric(n)>23)
royer_tax_over10 <- subset(royer_tax_count, as.numeric(n)>9)
```

Creating the linear models for comparing log lma and predictions for log(lma)
```{r}
royer_lmfam_top5 <- lm(log_lma~log_pet_leafarea, data = royer_tax_top5)
pred_lmfam_top5 <-  as.data.frame(predict(royer_lmfam_top5, interval="prediction"))
royer_lmfam_top5_bound <-  cbind(royer_tax_top5, pred_lmfam_top5)
```

Plotting log_lma vs log(petiole^2/leafarea) for the top 5 families
```{r}
ggplot(royer_lmfam_top5_bound, aes(log_pet_leafarea, log_lma))+
  aes(color=royer_lmfam_top5_bound$scrubbed_family)+
  geom_point() +
  geom_smooth(method=lm, se=TRUE)+
  labs(x="log(petiole^2/leafarea)")+
  labs(y="log(lma)")+
  labs(color = "Family")

ggplot(royer_lmfam_top5_bound, aes(log_pet_leafarea, log_lma))+
  geom_point() +
  geom_smooth(method=lm, se=TRUE)+
  labs(x="log(petiole^2/leafarea)")+
  labs(y="log(lma)")+
  labs(color = "Family")
```

###Logistic Regression
Logistic regression predicting phenology off of log_lma
```{r}
###Create sets so we can determine the top 5 extant families with the most samples, and the top 10 families respecitively

##Creation of fossil esitmates of top 5 fossil families
fossil_top5 <- royer_fossil_dropped%>% 
  group_by( scrubbed_family) %>% 
  tally()%>%
  filter(!grepl('Unknown', scrubbed_family))
fossil_top5 <- left_join(fossil_top5, royer_fossil_dropped, by="scrubbed_family")
fossil_top5 <- subset(fossil_top5, as.numeric(n)>22)

#Creation of extant off of the five top families
fossil_top5_tax <- as.data.frame(unique(fossil_top5$scrubbed_family))
extant_top5_fossil_tax <- subset(royer_tax_na_omit, (royer_tax_na_omit$scrubbed_family %in% fossil_top5_tax))

#creation of fossils that are in all extant families with more than 10 observations-----
extant_over10_fossil_tax <- subset(royer_tax_na_omit, (royer_tax_na_omit$scrubbed_family %in% royer_tax_over10$scrubbed_family))

##Create royer_data_split so genus and species are separated
royer_data<-read.csv("~/Documents/BEIN data R/data/raw/royer.data")
royer_data_split <-
  royer_data %>% separate(binomial, c("genus", "species"))
royer_data_split <- subset(royer_data_split, royer_data_split$genus %in% extant_over10_fossil_tax$scrubbed_genus)
royer_data_split <- subset(royer_data_split, royer_data_split$species %in% extant_over10_fossil_tax$species)
```

Logistic regression predictions on deciduous vs evergreen for fossils
```{r}
royer_data_split <- unique(royer_data_split)
colnames(royer_data_split)[colnames(royer_data_split)=="leaf_area"] <- "avg_LA"
glm_pred_data <- left_join(extant_over10_fossil_tax, royer_data_split, by = "avg_LA")
glm_pred_data <- na.omit(glm_pred_data)
glm_pred_data <- unique(glm_pred_data)


ggplot(glm_pred_data, aes(log_lma, phenology))+ 
  geom_point() + 
  stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)

ggplot(glm_pred_data, aes(log_lma, phenology, group=(scrubbed_family), color=scrubbed_family)) + 
  geom_point() + 
  stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
```