---
title: "LMA to Phenology"
output: html_document
---

```{r setup,}
require(BIEN)
require(plyr)
require(tidyr) 
require(dplyr)
require(mosaic)
require(stringr)
require(lme4)
require(magrittr)
require(data.table)
require(knitr)
require(kableExtra)
require(merTools)
require(boot)
require(cowplot)
require(ggridges)
require(tidyverse)
```
Loading processed data
```{r}
final_WSLA_DF<-readRDS(file="../data/processed/final_WSLA_DF.rds")
LMAtoPhenfossilpredictionsfinal<-readRDS(file="../data/processed/LMAtoPhenfossilpredictionsfinal.rds")
```

Model Creation and selection
```{r}
mod1<-glmer(Phenology2~log_LMA+(1+log_LMA|superorder/order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod2<-glmer(Phenology2~log_LMA+(1+log_LMA|order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod3<-glmer(Phenology2~log_LMA+(1+log_LMA|scrubbed_family), family="binomial", data=final_WSLA_DF)

mod4<-glmer(Phenology2~log_LMA+(0+log_LMA|superorder/order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod5<-glmer(Phenology2~log_LMA+(0+log_LMA|order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod6<-glmer(Phenology2~log_LMA+(0+log_LMA|scrubbed_family), family="binomial", data=final_WSLA_DF)

mod7<-glmer(Phenology2~log_LMA+(1|superorder/order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod8<-glmer(Phenology2~log_LMA+(1|order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod9<-glmer(Phenology2~log_LMA+(1|scrubbed_family), family="binomial", data=final_WSLA_DF)

mod10<-glmer(Phenology2~log_LMA+(1|scrubbed_family)+(1|order)+(1|superorder), family="binomial", data=final_WSLA_DF)
mod11<-glmer(Phenology2~log_LMA+(0+log_LMA|scrubbed_family)+(0+log_LMA|order)+(0+log_LMA|superorder), family="binomial", data=final_WSLA_DF)
mod12<-glmer(Phenology2~log_LMA+(1+log_LMA|scrubbed_family)+(1+log_LMA|order)+(1+log_LMA|superorder), family="binomial", data=final_WSLA_DF)

anova(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8, mod9, mod10, mod11, mod12)

ModTable<-data.frame("Model"=c(2, 12, 1, 3, 8, 10, 7, 9, 5, 11, 4, 6), "Rank"=c(1:12), "Slope Effects"=c("LLMA", "LLMA", "LLMA", "LLMA", "", "", "", "","LLMA", "LLMA", "LLMA", "LLMA" ), "Intercept Effects"=c("O, SF", "SO, O, SF", "SO, O, SF", "SF", "O, SF", "SO, O, SF", "SO, O, SF", "SF", "O, SF", "SO, O, SF","SO, O, SF", "SF"), "LogLik"=c(-10576, -10574, -10575, -10597, -12444, -12444, -12444, -12447, -12603, -12602, -12603, -12607), "AIC"=c(21169, 21171, 21171, 21204, 24896, 24897, 24898, 24899, 25214, 25215, 25216, 25220), "dAIC"=c(0, 2, 2, 35, 3727, 3728, 3729, 3730, 4045, 4046, 4047, 4051), "P-value"=c("<2e-16", "<2e-16", 0.3468, "<2e-16", "<2e-16", "<2e-16", "<2e-16", "<2e-16", 1, 1, 1, 1))

table<-kable(ModTable, align='c')%>%
  kable_styling(bootstrap_options="striped", full_width=F)
table

```

The best model is found to be mod2 with slope effect of log_LMA and intercept effects of order and scrubbed family. 
```{r}
mod2<-glmer(Phenology2~log_LMA+(1+log_LMA|order/scrubbed_family), family="binomial", data=final_WSLA_DF)
summary(mod2)
```

Prediction intervals are created for Phenology2 in fossilpredictionsfinal based on model 2. This gives prediction intervals for log LMA that give a more picture of what those values should be. 
```{r}
##Prediction Intervals 
PI<-predictInterval(merMod=mod2, newdata=LMAtoPhenfossilpredictionsfinal, 
                    level=0.95, n.sims=1000,
                    stat="median", type="linear.prediction",
                    include.resid.var = TRUE)
PI$fit<-inv.logit(PI$fit) 
PI$upr<-inv.logit(PI$upr)
PI$lwr<-inv.logit(PI$lwr)
finalpredictions<-cbind(LMAtoPhenfossilpredictionsfinal,PI)

ggplot(aes(x=log_LMA, y=fit, ymin=lwr, ymax=upr, color=finalpredictions$scrubbed_family), data=finalpredictions) +
  geom_point() +
  geom_linerange() +
  labs(color='scrubbed_family')

ggplot(finalpredictions, aes(log_LMA, fit, group=(scrubbed_family), color=scrubbed_family)) + 
  geom_point() + 
  stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)+
  xlab('log(Leaf Mass Area)')+
  ylab('Predicted Phenology')+
  labs(title = 'Predicted Phenology from LMA with Family Information')+
  labs(color='Family')
ggsave("LMAtoPhenFigure.pdf")
```

Joy Plot -- phenology vs log_lma with extant data
```{r}
library(ggplot2)
library(ggridges)

final_WSLA_DF<-readRDS(file="../data/processed/final_WSLA_DF.rds")

final_WSLA_DF_Unique<-unique(final_WSLA_DF)

final_WSLA_DF1<-subset(final_WSLA_DF_Unique, Phenology2==1)
final_WSLA_DF2<-subset(final_WSLA_DF_Unique, Phenology2==0)

intersect<-intersect(final_WSLA_DF1$scrubbed_family, final_WSLA_DF2$scrubbed_family)
intersect<-as.data.frame(intersect)
colnames(intersect)[colnames(intersect)=="intersect"]<-"scrubbed_family"
final_WSLA_DF<-final_WSLA_DF1[final_WSLA_DF$scrubbed_family %in% intersect$scrubbed_family,]

final_WSLA_DF1<-final_WSLA_DF1[final_WSLA_DF1$scrubbed_family %in% intersect$scrubbed_family,]
final_WSLA_DF2<-final_WSLA_DF2[final_WSLA_DF2$scrubbed_family %in% intersect$scrubbed_family,]

data<-rbind(final_WSLA_DF1, final_WSLA_DF2)
ggplot(data, aes(y=scrubbed_family,
                      x=log_LMA,
                      fill=as.factor(Phenology2))) +
  geom_density_ridges(scale=2) +
  scale_y_discrete(expand = c(0.01, 0)) +  
  scale_x_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c("salmon4", "darkgreen"), name="Phenology",
                         labels=c("Deciduous", "Evergreen"))+
  xlab('log(Leaf Mass Area)')+
  ylab('Family')+
  labs(title="LMA and Phenology of Extant Data")

ggsave("LMAtoPhenFigure2.pdf")
```
