---
title: "LMA to Phenology"
output: html_document
---

```{r setup,}
require(BIEN)
require(plyr)
require(tidyr) 
require(dplyr)
require(mosaic)
require(stringr)
require(lme4)
require(magrittr)
require(data.table)
require(knitr)
require(kableExtra)
require(merTools)
require(boot)
```


In order to run this model, we must first load royer_fossil_data_int and WSLA_final_DF. royer_fossil_data_int is our fossil dataset and WSLA_final_DF is our living dataset that will be used to create the prediction model. We must then eliminate unnecessary columns from each dataset and add a tally to the fossil set. 
```{r}
###final_WSLA_DF
final_WSLA_DF<-readRDS("03_final_WSLA_DF", file = "./data/processed/03_finalWSLADF.rds")
###Changing column names so they match up between files
colnames(final_WSLA_DF)[colnames(final_WSLA_DF)=="n"]<-"count"
###Removing columns that are irrelevant
final_WSLA_DF<-final_WSLA_DF[-c(2)]
```

Final fixes for final_WSLA_DF before model creation
```{r}
###assigns values to evergreen and deciduous and eliminates those without phenology 
final_WSLA_DF$Phenology<-as.factor(final_WSLA_DF$Phenology)
final_WSLA_DF$Phenology2<-ifelse(final_WSLA_DF$Phenology == "EV", 1, 0)
indx <- which(final_WSLA_DF$Phenology == "D_EV")
final_WSLA_DF$Phenology2[indx] <- NA

###Now we are going to add phenology and LMA back in so we have to match order and super order back up
species_df<-read.csv("./data/processed/species_df.csv")
final_WSLA_DF$binomial <- gsub("_", " ", final_WSLA_DF$binomial)
final_WSLA_DF<-left_join(species_df, final_WSLA_DF, by="binomial")
final_WSLA_DF$log_LMA<-log(final_WSLA_DF$LMA)
final_WSLA_DF<-final_WSLA_DF[-c(1, 5, 6, 8)]
final_WSLA_DF<-na.omit(final_WSLA_DF)
```

Model Creation and selection
```{r}
mod1<-glmer(Phenology2~log_LMA+(1+log_LMA|superorder/order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod2<-glmer(Phenology2~log_LMA+(1+log_LMA|order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod3<-glmer(Phenology2~log_LMA+(1+log_LMA|scrubbed_family), family="binomial", data=final_WSLA_DF)

mod4<-glmer(Phenology2~log_LMA+(0+log_LMA|superorder/order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod5<-glmer(Phenology2~log_LMA+(0+log_LMA|order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod6<-glmer(Phenology2~log_LMA+(0+log_LMA|scrubbed_family), family="binomial", data=final_WSLA_DF)

mod7<-glmer(Phenology2~log_LMA+(1|superorder/order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod8<-glmer(Phenology2~log_LMA+(1|order/scrubbed_family), family="binomial", data=final_WSLA_DF)
mod9<-glmer(Phenology2~log_LMA+(1|scrubbed_family), family="binomial", data=final_WSLA_DF)

mod10<-glmer(Phenology2~log_LMA+(1|scrubbed_family)+(1|order)+(1|superorder), family="binomial", data=final_WSLA_DF)
mod11<-glmer(Phenology2~log_LMA+(0+log_LMA|scrubbed_family)+(0+log_LMA|order)+(0+log_LMA|superorder), family="binomial", data=final_WSLA_DF)
mod12<-glmer(Phenology2~log_LMA+(1+log_LMA|scrubbed_family)+(1+log_LMA|order)+(1+log_LMA|superorder), family="binomial", data=final_WSLA_DF)

anova(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8, mod9, mod10, mod11, mod12)

ModTable<-data.frame("Model"=c(2, 12, 1, 3, 8, 10, 7, 9, 5, 11, 4, 6), "Rank"=c(1:12), "LogLik"=c(-10576, -10574, -10575, -10597, -12444, -12444, -12444, -12447, -12603, -12602, -12603, -12607), "AIC"=c(21169, 21171, 21171, 21204, 24896, 24897, 24898, 24899, 25214, 25215, 25216, 25220), "dAIC"=c(0, 2, 2, 35, 3727, 3728, 3729, 3730, 4045, 4046, 4047, 4051), "P-value"=c("<2e-16", "<2e-16", 0.3468, "<2e-16", "<2e-16", "<2e-16", "<2e-16", "<2e-16", 1, 1, 1, 1))

table<-kable(ModTable, align='c')%>%
  kable_styling(bootstrap_options="striped", full_width=F)
table

```

The best model is found to be mod2 with slope effect of log_LMA and intercept effects of order and scrubbed family. 
```{r}
mod2<-glmer(Phenology2~log_LMA+(1+log_LMA|order/scrubbed_family), family="binomial", data=final_WSLA_DF)
summary(mod2)
```

Moving forward, we must import the fossil data set with predicted log_lma values and eliminate columns that we do not need for the prediction model. 
```{r}
fossilpredictionsfinal<-read.csv("./data/processed/fossilpredictionsfinal.csv")
fossilpredictionsfinal<-fossilpredictionsfinal[c(5, 6, 8)]
colnames(fossilpredictionsfinal)[colnames(fossilpredictionsfinal)=="fit"]<- "log_LMA"
```

Prediction intervals are created for Phenology2 in fossilpredictionsfinal based on model 2. This gives prediction intervals for log LMA that give a more picture of what those values should be. 
```{r}
##Prediction Intervals 
PI<-predictInterval(merMod=mod2, newdata=fossilpredictionsfinal, 
                    level=0.95, n.sims=1000,
                    stat="median", type="linear.prediction",
                    include.resid.var = TRUE)
PI$fit<-inv.logit(PI$fit)
PI$upr<-inv.logit(PI$upr)
PI$lwr<-inv.logit(PI$lwr)
finalpredictions<-cbind(fossilpredictionsfinal,PI)

ggplot(aes(x=log_LMA, y=fit, ymin=lwr, ymax=upr, color=finalpredictions$scrubbed_family), data=finalpredictions) +
  geom_point() +
  geom_linerange() +
  labs(color='scrubbed_family')

ggplot(aes(x=log_LMA, y=fit, ymin=lwr, ymax=upr, color=finalpredictions$scrubbed_family), data=finalpredictions) +
  geom_point() +
  geom_linerange() +
  labs(color='scrubbed_family')

ggplot(finalpredictions, aes(log_LMA, fit, group=(scrubbed_family), color=scrubbed_family)) + 
  geom_point() + 
  stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)

```
