---
title: "Petiole and Leaf Area to LMA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Prep
Loading require packages
```{r}
require(BIEN)
require(plyr)
require(tidyr) 
require(dplyr)
require(mosaic)
require(stringr)
require(lme4)
require(magrittr)
```

First, we need to fix the fossil data necessary for prediction. Fossil data is compared to living data in order to come to conclusions when identifying the families of found fossils based off of LMA. The data has previously been adjusted to fix misspelled names, combine observations, and organize all of the information. 
```{r}
florissant_fossil_int <- readRDS("../data/processed/04_florissant_fossil_clean.rds")
renova_fossil_int <- readRDS("../data/processed/04_renova_fossil_clean.rds")
bridgecreek_fossil_int <- readRDS("../data/processed/04_bridgecreek_fossil_clean.rds")

fossil_comb <- rbind(florissant_fossil_int,renova_fossil_int,bridgecreek_fossil_int)
fossil_gen <- fossil_comb$Genus
fossil_gen <- unique(fossil_gen)
fossil_gen_df <- as.data.frame(fossil_gen)
fossil_tax <- BIEN_taxonomy_genus(fossil_gen)
fossil_tax <- fossil_tax[-c(1,7:9)]
fossil_tax <- unique(fossil_tax)

all_fossil <- rbind(florissant_fossil_int, renova_fossil_int, bridgecreek_fossil_int)
all_fossil$binomial <- paste(all_fossil$Genus, all_fossil$species)
all_fossil$binomial <- str_replace_all(all_fossil$binomial,"\\s+","_")
all_fossil <- all_fossil[-c(1,2)]
```

In the original data, species and genera are under 'binomial' and are presented as 'species_genus'. We must separate genera and species from one another so they can be analyzed separately.
```{r}
 all_fossil_LMEpred <- all_fossil %>%
  separate(binomial, 
    c("scrubbed_genus", "species"))
all_fossil_royer_pred <- left_join(all_fossil_LMEpred, fossil_tax, by = "scrubbed_genus")

all_fossil_royer_pred <- na.omit(all_fossil_royer_pred)
all_fossil_royer_pred <- unique(all_fossil_royer_pred)

colnames(all_fossil)[colnames(all_fossil)=="Petiole Width (cm)"] <- "avg_petiole_width"
colnames(all_fossil)[colnames(all_fossil)=="Leaf Area (cm^2)"] <- "avg_LA"
colnames(all_fossil)[colnames(all_fossil)=="PW^2/A"] <- "log_pet_leafarea"
colnames(all_fossil)[colnames(all_fossil)=="LMA (g/m^2)"] <- "log_LMA"
```


We first get rid of observations containing "na". Then, we create a leaf mass prediction and model using family as a factor. 
```{r}
royer_tax_full <- readRDS("../data/processed/07_lm4_royer")
royer_tax_na_omit<-na.omit(royer_tax_full)  

### Replacing "Unknown" with ""
royer_tax_na_omit$scrubbed_family[royer_tax_na_omit$scrubbed_family=="Unknown"] <- ""


tally<-tally(royer_tax_na_omit$scrubbed_family)
##Making the data able to be analyzed for slopes
  ###Adding occurence counts to dataset
royer_tax_count<-royer_tax_na_omit%>%
  group_by(scrubbed_family)%>%
  mutate(count=n())
royer_tax_count

  ###Removing observations of singular families n<3 and empty families
royer_tax_new<-subset(royer_tax_count, count>3 & scrubbed_family!="")


###Creating leaf mass prediction model
royer_lme <- lmer(log_lma~log_pet_leafarea + (1+log_pet_leafarea|scrubbed_family), data=royer_tax_new)

royer_lme_sum<- summary(royer_lme)
royer_lme_sum
```

Getting rid of observations in all_fossil_royer_pred that aren't present in royer_tax_new
```{r}
library(dataCompareR)
options(max.print = 1000)
af<-royer_tax_new$scrubbed_family
af
bf<-all_fossil_royer_pred$scrubbed_family
bf

##whats in all_fossil_royer_pred that isn't in #royer_tax_new

missing<-bf[!(bf%in%af)]
write.csv(missing, file="./data/missing.csv")

missing<-read.csv("./data/missing.csv")
missingdata<-subset(missing, select=-X)
colnames(missingdata)[colnames(missingdata)=="x"]<-"scrubbed_family"

tallymissing<-
  missingdata%>%
  group_by(scrubbed_family)%>%
  tally()
print.data.frame(tallymissing)

new_fossil_royer_pred2<-all_fossil_royer_pred[!all_fossil_royer_pred$scrubbed_family %in% missingdata$scrubbed_family,]

```

We use pattern matching and replacement to use the lm predictions and models using family as a factor.
```{r}
royer_fossil_dropped <-   all_fossil_royer_pred %>% 
  filter(!grepl('Cercidiphyllaceae', scrubbed_family)) %>%
  filter(!grepl('Smilacaceae', scrubbed_family)) %>%
  filter(!grepl('Staphyleaceae', scrubbed_family))
```

This code fixes royer_tax_full.
```{r}
royer_tax_fam_count <- royer_tax_full%>% 
  group_by(scrubbed_family) %>% 
  tally() %>%
  filter(!grepl('Unknown', scrubbed_family))

royer_tax_count <- left_join(royer_tax_full, royer_tax_fam_count, by="scrubbed_family")
royer_tax_new<-royer_tax_count
```

###Mixed Effects Modeling
Determining the top 5 extant families with the most samples, and the top 10 families respectively. 
```{r}
royer_tax_top5 <- subset(royer_tax_count, as.numeric(n)>23)
royer_tax_over10 <- subset(royer_tax_count, as.numeric(n)>9)
```

Creating royer_lmfam_top5 and royer_lm_top5, the linear models for comparing log lma and contains predictions for log(lma) for 5 fossil families. 
```{r}
royer_lmfam_top5 <- lm(log_lma~log_pet_leafarea, data = royer_tax_top5)
pred_lmfam_top5 <-  as.data.frame(predict(royer_lmfam_top5, interval="prediction"))
royer_lmfam_top5_bound <-  cbind(royer_tax_top5, pred_lmfam_top5)
```
non linear mixed effects model
-linear model on logit transform of response variable
-using non linear fitting and use link function link=logit prob lme4 can do it
-model comparison process will be the same, find the best by aic likelihood type analysis based on existing data then use to make predictions of fossil lmas
-model: none, family level, super order, order


Plotting log_lma vs log(petiole^2/leafarea) for the top 5 families. 
```{r}
ggplot(royer_lmfam_top5_bound, aes(log_pet_leafarea, log_lma))+
  aes(color=royer_lmfam_top5_bound$scrubbed_family)+
  geom_point() +
  geom_smooth(method=lm, se=TRUE)+
  labs(x="log(petiole^2/leafarea)")+
  labs(y="log(lma)")+
  labs(color = "Family")

ggplot(royer_lmfam_top5_bound, aes(log_pet_leafarea, log_lma))+
  geom_point() +
  geom_smooth(method=lm, se=TRUE)+
  labs(x="log(petiole^2/leafarea)")+
  labs(y="log(lma)")+
  labs(color = "Family")
```
