---
title: "Petiole Width to Leaf Mass Area"
output: html_document
---


The goal of this markdown is to show the paths taken to creating a model based off of a living data set that makes prediction intervals for LMA of a fossil data set. Once LMA is found, we will be able to gain phylogenetic information from a model that predicts phylogeny based off of LMA. This is being done to work on improving fossil species predictions when using taxonomic analysis of functional traits. 


```{r}
require(BIEN)
require(plyr)
require(tidyr) 
require(dplyr)
require(mosaic)
require(stringr)
require(lme4)
require(magrittr)
require(data.table)
require(knitr)
require(kableExtra)
```

First we must load all the processed data
```{r}
royer_tax_new<-readRDS(royer_tax_new, file="../data/processed/royer_tax_new.rds")
all_fossil_royer_pred<-readRDS(all_fossil_royer_pred, file="../data/processed/all_fossil_royer_pred.rds")
```

We must determine which mixed effects model is best suited to make prediction intervals. Using intercept effects super order, order, and scrubbed family and a slope effect of log pet leaf area, we determine the most suitable model by running a maximum likelihood ANOVA test and looking at the outputs in LogLik, AIC, dAIC, and p value. These are shown in the following table.
```{r}
model1<-lmer(log_lma~log_pet_leafarea + (1+log_pet_leafarea|superorder/order/scrubbed_family), data=royer_tax_new)
model2<-lmer(log_lma~log_pet_leafarea+(1+log_pet_leafarea|order/scrubbed_family), data=royer_tax_new)
model3<-lmer(log_lma~log_pet_leafarea + (1+log_pet_leafarea|scrubbed_family), data=royer_tax_new)
model4<-lmer(log_lma~log_pet_leafarea + (1|superorder/order/scrubbed_family), data=royer_tax_new)
model5<-lmer(log_lma~log_pet_leafarea + (1|order/scrubbed_family), data=royer_tax_new)
model6<-lmer(log_lma~log_pet_leafarea + (1|scrubbed_family), data=royer_tax_new)

modcompare<-anova(model1, model2, model3, model4, model5, model6)
ModelTable2<-data.frame("Model"=c(6, 3, 5, 2, 4, 1), "Rank"=c(1:6), "Slope Effects"=c("", "LPL","LPL", "", "LPL", ""), "Intercept Effects"=c("SF", "SF", "O, SF", "O, SF", "SO, O, SF", "SO, O, SF"), "LogLik"=c(-187.25, -184.00, -187.20, -183.69, -187.20, -183.32), "AIC"=c(382.49, 380.00, 384.40, 385.37, 386.40, 390.64), "dAIC"=c(0, -2.49, 4.40, 5.37, 6.40, 10.64))

table<-kable(ModelTable2, align='c')%>%
  kable_styling(bootstrap_options="striped", full_width=F)
table
```

The best model is found to be model 3, with a slope effect of log pet leaf area and intercept effect of scrubbed family.
```{r}
model3<-lmer(log_lma~log_pet_leafarea + (1+log_pet_leafarea|scrubbed_family), data=royer_tax_new)
```

In order to compare our fossil data set to the living data set, we must omit predictions from all_fossil_royer_pred that aren't present in royer_tax_new so all observations in the fossil data set are accounted for. 
```{r}
library(dataCompareR)
options(max.print = 1000)
af<-royer_tax_new$scrubbed_family
af
bf<-all_fossil_royer_pred$scrubbed_family
bf

##whats in all_fossil_royer_pred that isn't in #royer_tax_new

missing<-bf[!(bf%in%af)]
write.csv(missing, file="./data/missing.csv")

missing<-read.csv("./data/missing.csv")
missingdata<-subset(missing, select=-X)
colnames(missingdata)[colnames(missingdata)=="x"]<-"scrubbed_family"

tallymissing<-
  missingdata%>%
  group_by(scrubbed_family)%>%
  tally()
print.data.frame(tallymissing)

new_fossil_royer_pred2<-all_fossil_royer_pred[!all_fossil_royer_pred$scrubbed_family %in% missingdata$scrubbed_family,]
```

Then, prediction intervals are created for LMA in new_fossil_royer_pred2 based on model 3. The dataset contains observations for expected LMA, but this gives prediction intervals for log LMA that give a more picture of what those values should be. 
```{r}
##Prediction Intervals 
require(merTools)
PI<-predictInterval(merMod=model3, newdata=new_fossil_royer_pred2, 
                    level=0.95, n.sims=1000,
                    stat="median", type="linear.prediction",
                    include.resid.var = TRUE)
PI
fossilpredictions<-cbind(new_fossil_royer_pred2,PI)
library(ggplot2);
ggplot(aes(x=log_pet_leafarea, y=fit, ymin=lwr, ymax=upr, color=new_fossil_royer_pred2$scrubbed_family), data=fossilpredictions) +
  geom_point() +
  geom_linerange() +
  labs(color='scrubbed_family')
```

Save fossil predictions and eliminate unnecessary columns
```{r}
fossilpredictionsfinal<-fossilpredictions[-c(2, 4, 5, 6, 7, 8)]
write.csv(fossilpredictionsfinal, file="./data/processed/fossilpredictionsfinal.csv")
```
